Turnbull, C. M. (1972). The mountain people. Touchstone Book

Alburez-Gutierrez, D., Kolk, M., & Zagheni, E. (2021). Women’s experience of child death over the life course: A global demographic perspective. Demography, Advanced publication.

Smith-Greenaway, E., Alburez-Gutierrez, D., Trinitapoli, J., & Zagheni, E. (2021). Global burden of maternal bereavement: Indicators of the cumulative prevalence of child loss. BMJ Global Health, 6(4), e004837

Alburez-Gutierrez, D., Mason, C., & Zagheni, E. (2021). The “Sandwich Generation” Revisited: Global Demographic Drivers of Care Time Demands. Population and Development Review, Advanced publication.

Alburez-Gutierrez, D. (2019). Blood is thicker than bloodshed: A genealogical approach to reconstruct populations after armed conﬂicts. Demographic Research, 40, 627–656.

---------

\footnote{Turnbull, C. M. (1972). The mountain people. Touchstone Book}

\footfullcite{alburezgutierrez_sandwich_2021}

\footnote{Alburez-Gutierrez, D., et al. (2021). ``The `Sandwich Generation' Revisited: Global Demographic Drivers of Care Time Demands.'' \textbf{Population and Development Review}. Advanced Publication.}

\footnote{Alburez-Gutierrez, , et al. (2021) ``Women's experience of child death: A global demographic perspective.'' \textbf{Demography}. Advanced Publication.}

\footfullcite{smith-greenaway_global_2021}

\footnote{Smith-Greenaway, E., Alburez-Gutierrez, D., et al. (2021) ``The Global Burden of Maternal Bereavement: Indicators of the Cumulative Prevalence of Child Loss.'' \textbf{BMJ Global Health}, 6:e004837.}

\footnote{Martinez-Folgar, K., Alburez-Gutierrez, D., et al. (2021). ``Excess Mortality during the COVID-19 Pandemic in Guatemala''. \textbf{American Journal of Public Health}, 0:e1-e8.}

\footnote{Nepomuceno, M., Acosta, E., Alburez-Gutierrez, D., et al. (2020). ``Besides population age-structure, health and other demographic factors can contribute to understanding the COVID-19 burden." \textbf{Proceedings of the National Academy of Sciences}, 117(25).}


\footfullcite{alburez-gutierrez_blood_2019}

\footnote{Alburez-Gutierrez, D. (2019). Blood is thicker than bloodshed: A genealogical approach to reconstruct populations after armed conflicts. \textbf{Demographic Research}, 40(23): 627-656.}

\textbf{Alburez-Gutierrez, D.} (2021). ``The demographic drivers of grief and memory after genocide in Guatemala.'' \textbf{R \& R in Demography}.
